*DCVtestkit*: a R package for linearity assessment and visualisation of multiple curves

### 2022-11-09 Manuscript/manuscript.qmd
Fixed equation alignment issue and changed title.

---
title: "*DCVtestkit*: a R package for linearity assessment and analysis of quality control dilution curves"

@@

Next, summary statistics are calculated via `summarise_dilution_table` for each dilution curve. Besides the Pearson Correlation Coefficient, one additional calculation is the Mandel's Fitting Test (@eq-mandel-test) from @andrade2013, in which $\mathrm S_{y/x,lin}^2$ and $\mathrm S_{y/x,quad}^2$ are the standard error of the regression divided by the regression's degrees of freedom for the linear model (@eq-s1) and quadratic model (@eq-s2) repsectively. A low $p$ value from the $F$ test gives sufficient evidence that a quadratic model fits better than a linear model, indicating the curve may not be linear.

$$
\mathrm F_{test} = \frac{(\mathrm n-2) \ \mathrm S_{y/x,lin}^2 \ - \ (\mathrm n-3) \ \mathrm S_{y/x,quad}^2}{\mathrm S_{y/x,quad}^2}
$$ {#eq-mandel-test}

$$
\mathrm S_{y/x,lin}^2 = \frac{1}{n-2} \sum_{i=1}^{n} ( y_{i,pred} - y_{i,true})^2
$$ {#eq-s1}

$$
\mathrm S_{y/x,quad}^2 = \frac{1}{n-3} \sum_{i=1}^{n} ( y_{i,pred} - y_{i,true})^2
$$ {#eq-s2}

Another one is Percent Residual Accuracy (@eq-pra) from @logue2018 in which $x_{i,true}$ is the true x-axis value and $x_{i,calc}$ is the predicted x-axis value from the linear model. Ranging from $-\infty$ to $100$, if the curve is linear, the value should be close to $100$.

$$
\% \ RA = \frac{100\%}{n} \sum_{i=1}^{n}\left(1 - \bigg| \frac{x_{i,true}-x_{i,pred}}{x_{i,true}} \bigg|\right)
$$ {#eq-pra}

>>>>

---
title: "*DCVtestkit*: a R package for linearity assessment and visualisation of multiple curves"

@@

Next, summary statistics are calculated via `summarise_dilution_table` for each dilution curve. Besides the Pearson Correlation Coefficient, one additional calculation is the Mandel's Fitting Test ($F_{stats}$ in @eq-mandel-test) from @andrade2013. A low $p$ value from the $F$ test gives sufficient evidence that a quadratic model fits better than a linear model, indicating the curve may not be linear.

$$
\begin{aligned}
&F_{stats} = \frac{(n-2) \ \mathrm S_{lin}^2 \ - \ (n-3) \ \mathrm S_{quad}^2}{\mathrm S_{quad}^2} \sim \ F(1,n-3)
\\
\\
&\text{where}
\\
&\mathrm S_{lin}^2 = \frac{1}{n-2} \sum_{i=1}^{n} ( y_{i,lin} - y_{i,true})^2 
\\
&\mathrm S_{quad}^2 = \frac{1}{n-3} \sum_{i=1}^{n} ( y_{i,quad} - y_{i,true})^2
\\ 
&y_{i,lin} \ \text{is the linear model predicted y-axis value.}
\\ 
&y_{i,quad} \ \text{is the quadratic model predicted y-axis value.}
\\ 
&y_{i,true} \ \text{is the true y-axis value.}
\\ 
&n \ \text{is the number of data points.}
\end{aligned}
$$ {#eq-mandel-test}
Another one is Percent Residual Accuracy ($\%RA$ in @eq-pra) from @logue2018. Ranging from $-\infty$ to $100$, if the curve is linear, the value should be close to $100$.

$$
\begin{aligned}
&\%RA = \frac{100\%}{n} \sum_{i=1}^{n}\left(1 - \bigg| \frac{x_{i,true}-x_{i,lin}}{x_{i,true}} \bigg|\right)
\\
\\
&\text{where}
\\
&x_{i,true} \ \text{is the true x-axis value.}
\\ 
&x_{i,lin} \ \text{is the linear model predicted x-axis value.}
\\ 
&n \ \text{is the number of data points.}
\end{aligned}
$$ {#eq-pra}
